---
title: Klasse im Puls
author: 
  - Christoph Wunder
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    fig_caption: yes
    theme: spacelab 
    highlight: pygments
    toc: TRUE
    toc_depth: 2
    number_sections: TRUE
    toc_float:
      smooth_scroll: FALSE
bibliography: literature.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment=NA)
rm(list=ls())
```
```{r, include=TRUE, message = FALSE}
library(tidyverse)
library(Hmisc)
library(mosaic)
library(foreign)
library(lme4)
library(rstanarm)
library(brms)
library(tableone)
library(RColorBrewer)
library(xtable)
library(knitr)
library(kableExtra)
library(ggdist)
library(plm)
```

```{r}
table(d2$female, d2$school, useNA = "ifany")
```


```{r, message=FALSE, cache=TRUE}
bayes4 <- stan_lmer( lsat ~ treated + welle + tXw2 + female + female:tXw2 +
                       (1 + treated + tXw2 + female + female:tXw2 | school) + (1 | pid), 
                data=d2, chains = 4, cores = 4, seed=123 )
summary(bayes4, 
        pars = c("(Intercept)", "treated1", "welle2", "tXw21", "female1", "tXw21:female1",
                 "b[tXw21 school:1]", "b[tXw21 school:2]", "b[tXw21 school:3]",
                 "b[tXw21 school:4]", "b[tXw21 school:5]", "b[tXw21 school:6]",
                 "b[tXw21:female1 school:1]", "b[tXw21:female1 school:2]", "b[tXw21:female1 school:3]",
                 "b[tXw21:female1 school:4]", "b[tXw21:female1 school:5]", "b[tXw21:female1 school:6]"),
        probs = c(0.1, 0.5, 0.9),
        digits = 4)
```

```{r}
teff2_gender <- as.matrix(bayes4, 
                   pars = c("tXw21", "tXw21:female1",
                            "b[tXw21 school:1]", "b[tXw21:female1 school:1]",
                            "b[tXw21 school:2]", "b[tXw21:female1 school:2]",
                            "b[tXw21 school:3]", "b[tXw21:female1 school:3]",
                            "b[tXw21 school:4]", "b[tXw21:female1 school:4]",
                            "b[tXw21 school:5]", "b[tXw21:female1 school:5]",
                            "b[tXw21 school:6]", "b[tXw21:female1 school:6]")) 

school1_boys <- teff2_gender[,"tXw21"] + teff2_gender[,"b[tXw21 school:1]"]
school2_boys <- teff2_gender[,"tXw21"] + teff2_gender[,"b[tXw21 school:2]"]
school3_boys <- teff2_gender[,"tXw21"] + teff2_gender[,"b[tXw21 school:3]"]
school4_boys <- teff2_gender[,"tXw21"] + teff2_gender[,"b[tXw21 school:4]"]
school5_boys <- teff2_gender[,"tXw21"] + teff2_gender[,"b[tXw21 school:5]"]
school6_boys <- teff2_gender[,"tXw21"] + teff2_gender[,"b[tXw21 school:6]"]
school1_girls <- teff2_gender[,"tXw21"] + teff2_gender[,"tXw21:female1"] + teff2_gender[,"b[tXw21 school:1]"] + teff2_gender[,"b[tXw21:female1 school:1]"]
school2_girls <- teff2_gender[,"tXw21"] + teff2_gender[,"tXw21:female1"] + teff2_gender[,"b[tXw21 school:2]"] + teff2_gender[,"b[tXw21:female1 school:2]"]
school3_girls <- teff2_gender[,"tXw21"] + teff2_gender[,"tXw21:female1"] + teff2_gender[,"b[tXw21 school:3]"] + teff2_gender[,"b[tXw21:female1 school:3]"]
school4_girls <- teff2_gender[,"tXw21"] + teff2_gender[,"tXw21:female1"] + teff2_gender[,"b[tXw21 school:4]"] + teff2_gender[,"b[tXw21:female1 school:4]"]
school5_girls <- teff2_gender[,"tXw21"] + teff2_gender[,"tXw21:female1"] + teff2_gender[,"b[tXw21 school:5]"] + teff2_gender[,"b[tXw21:female1 school:5]"]
school6_girls <- teff2_gender[,"tXw21"] + teff2_gender[,"tXw21:female1"] + teff2_gender[,"b[tXw21 school:6]"] + teff2_gender[,"b[tXw21:female1 school:6]"]

teff2_boys <- data.frame( 
                     school1 = school1_boys,
                     school2 = school2_boys,
                     school3 = school3_boys,
                     school4 = school4_boys,
                     school5 = school5_boys,
                     school6 = school6_boys)

teff2_girls <- data.frame( 
                     school1 = school1_girls,
                     school2 = school2_girls,
                     school3 = school3_girls,
                     school4 = school4_girls,
                     school5 = school5_girls,
                     school6 = school6_girls)

```

```{r}
r2_mean_boys <- apply( X = teff2_boys, MARGIN = 2, FUN = mean )
r2_mean_girls <- apply( X = teff2_girls, MARGIN = 2, FUN = mean )
r2_quant_boys <- apply( X = teff2_boys, MARGIN = 2, FUN = quantile, probs = c(0.1, 0.9))
r2_quant_girls <- apply( X = teff2_girls, MARGIN = 2, FUN = quantile, probs = c(0.1, 0.9))
r2_quant_boys <- data.frame(t(r2_quant_boys))
r2_quant_girls <- data.frame(t(r2_quant_girls))
names(r2_quant_boys) <- c("Q10", "Q90")
names(r2_quant_girls) <- c("Q10", "Q90")
r2_boys <- data.frame( school = seq(1:6), mean = r2_mean_boys, r2_quant_boys, wave = 2, female = as.factor(0) )
r2_girls <- data.frame( school = seq(1:6), mean = r2_mean_girls, r2_quant_girls, wave = 2, female = as.factor(1) )
r2_gender <- rbind.data.frame(r2_boys, r2_girls)
r2_gender
```

```{r}
ggplot(data = r2_gender, 
       aes(x = wave, 
           y = mean)) +
  geom_pointrange(aes(ymin = Q10, 
                      ymax = Q90, group = school, color = school),
                  position=position_dodge(width=0.20)) +
  scale_x_discrete("Wave", limits = c(2)) +
  geom_hline(yintercept=0) + facet_grid(~ female)
```


# Alternative model: varying intercepts for time trend across schools 

## Model statement

We allow for different time trends across schools. This preserves the common trend assumption within schools. The regression equation is:
$$
\begin{align}
y_{ijt} = \beta_0 + \beta_{1j[i]} T_i + \sum\limits_{l=2}^5\delta_{lj[i]} (T_{i} \times \text{Period}_{t}) + \lambda_{tj[i]} + \mu_{j[i]} + \alpha_{i} + \epsilon_{ijt},
\end{align}
$$

# Individual fixed effects

## Model statement

$$
\begin{align}
y_{ijt} = \beta_0 + \beta_{1j[i]} T_i + \sum\limits_{l=2}^5\delta_{lj[i]} (T_{i} \times \text{Period}_{t}) + \lambda_{t}[i] + \mu_{j[i]} + \alpha_{i} + \epsilon_{ijt},
\end{align}
$$

* $\alpha_{i}$ models individual-specific unobserved factors. Individual heterogeneity is model because we have repeated observations for the same pupils over time. We do not specify a model the individual-specific term $\alpha_i$ here. This is equivalent to using dummy variables for all pupils (and omitting the overall intercept).

# Original computing environment

```{r, results='hide'}
devtools::session_info()
```

