---
title: Klasse im Puls
author: 
  - Christoph Wunder
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    fig_caption: yes
    theme: spacelab 
    highlight: pygments
    toc: TRUE
    toc_depth: 2
    number_sections: TRUE
    toc_float:
      smooth_scroll: FALSE
bibliography: literature.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment=NA)
rm(list=ls())
```
```{r, include=TRUE, message = FALSE}
library(tidyverse)
library(Hmisc)
library(mosaic)
library(foreign)
library(lme4)
library(rstanarm)
library(brms)
library(tableone)
library(RColorBrewer)
library(xtable)
library(knitr)
library(kableExtra)
library(ggdist)
library(plm)
```
# Alternative model: varying intercepts for time trend across schools 

## Model statement

We allow for different time trends across schools. This preserves the common trend assumption within schools. The regression equation is:
$$
\begin{align}
y_{ijt} = \beta_0 + \beta_{1j[i]} T_i + \sum\limits_{l=2}^5\delta_{lj[i]} (T_{i} \times \text{Period}_{t}) + \lambda_{tj[i]} + \mu_{j[i]} + \alpha_{i} + \epsilon_{ijt},
\end{align}
$$

# Individual fixed effects

## Model statement

$$
\begin{align}
y_{ijt} = \beta_0 + \beta_{1j[i]} T_i + \sum\limits_{l=2}^5\delta_{lj[i]} (T_{i} \times \text{Period}_{t}) + \lambda_{t}[i] + \mu_{j[i]} + \alpha_{i} + \epsilon_{ijt},
\end{align}
$$

* $\alpha_{i}$ models individual-specific unobserved factors. Individual heterogeneity is model because we have repeated observations for the same pupils over time. We do not specify a model the individual-specific term $\alpha_i$ here. This is equivalent to using dummy variables for all pupils (and omitting the overall intercept).

# Original computing environment

```{r, results='hide'}
devtools::session_info()
```

