---
title: | 
       | Klasse im Puls
       | Additional outcomes: psychological outcomes 
author: 
  - Christoph Wunder
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    fig_caption: yes
    theme: spacelab 
    highlight: pygments
    toc: TRUE
    toc_depth: 2
    number_sections: TRUE
    toc_float:
      smooth_scroll: FALSE
bibliography: "../paper/literature.bib"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment=NA)
rm(list=ls())
```
```{r, include=TRUE, message = FALSE}
library(tidyverse)
library(Hmisc)
library(mosaic)
library(foreign)
library(lme4)
library(rstanarm)
library(brms)
library(tableone)
library(RColorBrewer)
library(xtable)
library(knitr)
library(kableExtra)
```


# Data

```{r}
d <- readRDS(file = "../data/d4.RDS")
```

```{r}
summary(d)
```



# Standardized big 5 index 

## Model statement

$$
\begin{align}
y_{ijt} = \beta_0 + \beta_{1j[i]} T_i + \sum\limits_{l=2}^5\delta_{lj[i]} (T_{i} \times \text{Period}_{t}) + \lambda_t + \mu_{j[i]} + \alpha_{i} + \epsilon_{ijt},
\end{align}
$$

## Estimation

```{r, cache=TRUE}
prior <- c(prior(normal(0, 5), class = Intercept),
           prior(normal(0, 5), class = b),
           prior(cauchy(0, 1), class = sd))
m1 <- brm( std_big5index ~ 
             treated + welle + tXw2 + tXw3 + tXw4 + tXw5 + 
             female + migback +
             (1 + treated + tXw2 + tXw3 + tXw4 + tXw5 | school) + (1 | pid),
           prior = prior, cores = 4, chains = 4, 
           data = d, seed=123,
           file = "../results/psych_m1")
print(m1)
```

## Posterior predictions for average (or marginal) school

```{r}
x_pred_avg <- tibble( treated = as.factor( rep(c(0, 1), 5)),
                  welle   = as.factor( c(1, 1, 2, 2, 3, 3, 4, 4, 5, 5) ),
                  tXw2    = c(0,0,0,1,0,0,0,0,0,0),
                  tXw3    = c(0,0,0,0,0,1,0,0,0,0),
                  tXw4    = c(0,0,0,0,0,0,0,1,0,0),
                  tXw5    = c(0,0,0,0,0,0,0,0,0,1),
                  female  = as.factor( rep(1, 10) ),
                  migback = as.factor( rep("Native", 10) ))
y_pred <- fitted( m1, newdata = x_pred_avg, 
                  re_formula = NA,
                  probs = c(0.25, 0.75) )
pred_df <- cbind.data.frame( y_pred, x_pred_avg )
```

```{r}
gr_std_big5index_pred <-  ggplot( pred_df, aes( x = welle, y = Estimate, group = treated) ) + 
  geom_line( aes(color=treated) ) +
  geom_point( aes(color=treated) ) +
  geom_pointrange(aes(ymin = Q25, 
                      ymax = Q75, color = treated ),
                  position=position_dodge(width=0.1) ) +
  ylab("Std. Big 5 Index") +
  scale_x_discrete("Wave", limits = seq(1:5) ) +
  theme(legend.position="bottom")
ggsave("../figures/std_big5index_pred.pdf", gr_std_big5index_pred, width = 7, height = 4, units = "in")
gr_std_big5index_pred
```

## Average treatment effect

```{r}
post <- posterior_samples(m1)
teff <- tibble( teff2 = post$`b_tXw2`,
                teff3 = post$`b_tXw3`, 
                teff4 = post$`b_tXw4`, 
                teff5 = post$`b_tXw5` )
teff_mean <- apply( X = teff, MARGIN = 2, FUN = mean )
teff_quant <- t(apply( X = teff, MARGIN = 2, 
                     FUN = quantile, probs = c(0.1, 0.9) ))
su_teff <- tibble( mean = teff_mean, 
                   Q10 = teff_quant[,1],
                   Q90 = teff_quant[,2], 
                   wave = seq(2:5) )
su_teff
```

```{r} 
gr_std_big5index_teff <- ggplot(data = su_teff, 
       aes(x = wave, 
           y = mean)) +
  geom_pointrange(aes(ymin = Q10, ymax = Q90)) +
  scale_x_discrete("Wave", limits=c("2", "3", "4", "5")) +
  geom_hline(yintercept=0) +
  ylab("Average treatment effect")
ggsave("../figures/std_big5index_teff.pdf", gr_std_big5index_teff, width = 7, height = 4, units = "in")
gr_std_big5index_teff
```

## Posterior predictions for existing schools

```{r}
x_pred_schools <- tibble( treated = as.factor( rep(rep(c(0, 1), 5), 6) ),
                  welle   = as.factor( rep(rep(1:5, each=2), 6) ),
                  school  = as.factor( rep(1:6, each=10) ),
                  tXw2    = rep(c(0,0,0,1,0,0,0,0,0,0), 6),
                  tXw3    = rep(c(0,0,0,0,0,1,0,0,0,0), 6),
                  tXw4    = rep(c(0,0,0,0,0,0,0,1,0,0), 6),
                  tXw5    = rep(c(0,0,0,0,0,0,0,0,0,1), 6),
                  female  = as.factor( rep(1, 60) ),
                  migback = as.factor( rep("Native", 60) ))
y_pred <- fitted( m1, newdata = x_pred_schools, 
                  re_formula = ~(1 + treated + tXw2 + tXw3 + tXw4 + tXw5 | school) ) 
pred_df <- cbind.data.frame( y_pred, x_pred_schools )

gr_std_big5index_pred_across_schools <- 
  ggplot( pred_df, aes( x = welle, y = Estimate, group = treated) ) + 
  geom_line( aes(color=treated) ) +
  ylab("Std. Big 5 Index") +
  scale_x_discrete("Wave", limits = seq(1:5) ) +
  facet_wrap(vars(school), labeller = labeller(
    school = c("1" = "School: 1", "2" = "School: 2", "3" = "School: 3",
               "4" = "School: 4", "5" = "School: 5", "6" = "School: 6" ))) +
  theme(legend.position="bottom")
ggsave("../figures/std_big5index_pred_across_schools.pdf", gr_std_big5index_pred_across_schools, 
       width = 7, height = 4, units = "in")
gr_std_big5index_pred_across_schools
```

## Treatment effect heterogeneity across schools

```{r}
post <- posterior_samples(m1)
teff_w2 <- tibble( school1 = post$`b_tXw2` + post$`r_school[1,tXw21]`,
                   school2 = post$`b_tXw2` + post$`r_school[2,tXw21]`,
                   school3 = post$`b_tXw2` + post$`r_school[3,tXw21]`,
                   school4 = post$`b_tXw2` + post$`r_school[4,tXw21]`,
                   school5 = post$`b_tXw2` + post$`r_school[5,tXw21]`,
                   school6 = post$`b_tXw2` + post$`r_school[6,tXw21]` )
teff_mean_w2 <- apply( X = teff_w2, MARGIN = 2, FUN = mean )
teff_quant_w2 <- t(apply( X = teff_w2, MARGIN = 2, 
                     FUN = quantile, probs = c(0.1, 0.9) ))

teff_w3 <- tibble( school1 = post$`b_tXw3` + post$`r_school[1,tXw31]`,
                   school2 = post$`b_tXw3` + post$`r_school[2,tXw31]`,
                   school3 = post$`b_tXw3` + post$`r_school[3,tXw31]`,
                   school4 = post$`b_tXw3` + post$`r_school[4,tXw31]`,
                   school5 = post$`b_tXw3` + post$`r_school[5,tXw31]`,
                   school6 = post$`b_tXw3` + post$`r_school[6,tXw31]` )
teff_mean_w3 <- apply( X = teff_w3, MARGIN = 2, FUN = mean )
teff_quant_w3 <- t(apply( X = teff_w3, MARGIN = 2, 
                     FUN = quantile, probs = c(0.1, 0.9) ))

teff_w4 <- tibble( school1 = post$`b_tXw4` + post$`r_school[1,tXw41]`,
                   school2 = post$`b_tXw4` + post$`r_school[2,tXw41]`,
                   school3 = post$`b_tXw4` + post$`r_school[3,tXw41]`,
                   school4 = post$`b_tXw4` + post$`r_school[4,tXw41]`,
                   school5 = post$`b_tXw4` + post$`r_school[5,tXw41]`,
                   school6 = post$`b_tXw4` + post$`r_school[6,tXw41]` )
teff_mean_w4 <- apply( X = teff_w4, MARGIN = 2, FUN = mean )
teff_quant_w4 <- t(apply( X = teff_w4, MARGIN = 2, 
                     FUN = quantile, probs = c(0.1, 0.9) ))

teff_w5 <- tibble( school1 = post$`b_tXw5` + post$`r_school[1,tXw51]`,
                   school2 = post$`b_tXw5` + post$`r_school[2,tXw51]`,
                   school3 = post$`b_tXw5` + post$`r_school[3,tXw51]`,
                   school4 = post$`b_tXw5` + post$`r_school[4,tXw51]`,
                   school5 = post$`b_tXw5` + post$`r_school[5,tXw51]`,
                   school6 = post$`b_tXw5` + post$`r_school[6,tXw51]` )
teff_mean_w5 <- apply( X = teff_w5, MARGIN = 2, FUN = mean )
teff_quant_w5 <- t(apply( X = teff_w5, MARGIN = 2, 
                     FUN = quantile, probs = c(0.1, 0.9) ))

su_teff <- tibble( school = rep(seq(1:6), 4), 
                   mean = c(teff_mean_w2, teff_mean_w3,
                            teff_mean_w4, teff_mean_w5), 
                   Q10 = c(teff_quant_w2[,1], teff_quant_w3[,1],
                           teff_quant_w4[,1], teff_quant_w5[,1]),
                   Q90 = c(teff_quant_w2[,2], teff_quant_w3[,2],
                           teff_quant_w4[,2], teff_quant_w5[,2]),
                   wave = rep(2:5, each=6) )
```
Graphical representation of results.

```{r}
gr_std_big5index_teff_across_schools <-
  ggplot(data = su_teff, aes(x = wave, y = mean)) +
  geom_pointrange(aes(ymin = Q10, 
                      ymax = Q90, group = school, color = school),
                  position=position_dodge(width=0.20)) +
  scale_x_discrete("Wave", limits=seq(2:6) )  +
  geom_hline(yintercept=0) + 
  theme(legend.position="bottom")
ggsave("../figures/std_big5index_teff_across_schools.pdf", gr_std_big5index_teff_across_schools, 
       width = 7, height = 4, units = "in")
gr_std_big5index_teff_across_schools
```

# Standardized emotional stability  

## Estimation

```{r, cache=TRUE}
prior <- c(prior(normal(0, 5), class = Intercept),
           prior(normal(0, 5), class = b),
           prior(cauchy(0, 1), class = sd))
m2 <- brm( std_emostab ~ treated + welle + tXw2 + tXw3 + tXw4 + tXw5 + female + migback +
             (1 + treated + tXw2 + tXw3 + tXw4 + tXw5 | school) + (1 | pid),
           prior = prior, cores = 4, chains = 4, 
           data = d, seed=123,
           file = "../results/psych_m2")
print(m2)
```

## Posterior predictions for average (or marginal) school

```{r}
y_pred <- fitted( m2, newdata = x_pred_avg, 
                  re_formula = NA,
                  probs = c(0.25, 0.75) )
pred_df <- cbind.data.frame( y_pred, x_pred_avg )
```

```{r}
gr_std_emostab_pred <-  ggplot( pred_df, aes( x = welle, y = Estimate, group = treated) ) + 
  geom_line( aes(color=treated) ) +
  geom_point( aes(color=treated) ) +
  geom_pointrange(aes(ymin = Q25, 
                      ymax = Q75, color = treated ),
                  position=position_dodge(width=0.1) ) +
  ylab("Std. Emotional Stability") +
  scale_x_discrete("Wave", limits = seq(1:5) ) +
  theme(legend.position="bottom")
ggsave("../figures/std_emostab_pred.pdf", gr_std_emostab_pred, width = 7, height = 4, units = "in")
gr_std_emostab_pred
```

## Average treatment effect

```{r}
post <- posterior_samples(m2)
teff <- tibble( teff2 = post$`b_tXw2`,
                teff3 = post$`b_tXw3`, 
                teff4 = post$`b_tXw4`, 
                teff5 = post$`b_tXw5` )
teff_mean <- apply( X = teff, MARGIN = 2, FUN = mean )
teff_quant <- t(apply( X = teff, MARGIN = 2, 
                     FUN = quantile, probs = c(0.1, 0.9) ))
su_teff <- tibble( mean = teff_mean, 
                   Q10 = teff_quant[,1],
                   Q90 = teff_quant[,2], 
                   wave = seq(2:5) )
su_teff
```

```{r} 
gr_std_emostab_teff <- ggplot(data = su_teff, 
       aes(x = wave, 
           y = mean)) +
  geom_pointrange(aes(ymin = Q10, ymax = Q90)) +
  scale_x_discrete("Wave", limits=c("2", "3", "4", "5")) +
  geom_hline(yintercept=0) +
  ylab("Average treatment effect")
ggsave("../figures/std_big5index_teff.pdf", gr_std_emostab_teff, width = 7, height = 4, units = "in")
gr_std_emostab_teff
```

## Posterior predictions for existing schools

```{r}
x_pred <- tibble( treated = as.factor( rep(rep(c(0, 1), 5), 6) ),
                  welle   = as.factor( rep(rep(1:5, each=2), 6) ),
                  school  = as.factor( rep(1:6, each=10) ),
                  tXw2    = rep(c(0,0,0,1,0,0,0,0,0,0), 6),
                  tXw3    = rep(c(0,0,0,0,0,1,0,0,0,0), 6),
                  tXw4    = rep(c(0,0,0,0,0,0,0,1,0,0), 6),
                  tXw5    = rep(c(0,0,0,0,0,0,0,0,0,1), 6),
                  female  = as.factor( rep(1, 60) ),
                  migback = as.factor( rep("Native", 60) ))
y_pred <- fitted( m2, newdata = x_pred, 
                  re_formula = ~(1 + treated + tXw2 + tXw3 +
                                   tXw4 + tXw5 | school) ) 
pred_df <- cbind.data.frame( y_pred, x_pred )

gr_std_emostab_pred_across_schools <- 
  ggplot( pred_df, aes( x = welle, y = Estimate, group = treated) ) + 
  geom_line( aes(color=treated) ) +
  ylab("Std. Big 5 Index") +
  scale_x_discrete("Wave", limits = seq(1:5) ) +
  facet_wrap(vars(school), labeller = labeller(
    school = c("1" = "School: 1", "2" = "School: 2", "3" = "School: 3",
               "4" = "School: 4", "5" = "School: 5", "6" = "School: 6" ))) +
  theme(legend.position="bottom")
ggsave("../figures/std_emostab_pred_across_schools.pdf", gr_std_emostab_pred_across_schools, 
       width = 7, height = 4, units = "in")
gr_std_emostab_pred_across_schools
```

## Treatment effect heterogeneity across schools

```{r}
post <- posterior_samples(m2)
teff_w2 <- tibble( school1 = post$`b_tXw2` + post$`r_school[1,tXw21]`,
                   school2 = post$`b_tXw2` + post$`r_school[2,tXw21]`,
                   school3 = post$`b_tXw2` + post$`r_school[3,tXw21]`,
                   school4 = post$`b_tXw2` + post$`r_school[4,tXw21]`,
                   school5 = post$`b_tXw2` + post$`r_school[5,tXw21]`,
                   school6 = post$`b_tXw2` + post$`r_school[6,tXw21]` )
teff_mean_w2 <- apply( X = teff_w2, MARGIN = 2, FUN = mean )
teff_quant_w2 <- t(apply( X = teff_w2, MARGIN = 2, 
                     FUN = quantile, probs = c(0.1, 0.9) ))

teff_w3 <- tibble( school1 = post$`b_tXw3` + post$`r_school[1,tXw31]`,
                   school2 = post$`b_tXw3` + post$`r_school[2,tXw31]`,
                   school3 = post$`b_tXw3` + post$`r_school[3,tXw31]`,
                   school4 = post$`b_tXw3` + post$`r_school[4,tXw31]`,
                   school5 = post$`b_tXw3` + post$`r_school[5,tXw31]`,
                   school6 = post$`b_tXw3` + post$`r_school[6,tXw31]` )
teff_mean_w3 <- apply( X = teff_w3, MARGIN = 2, FUN = mean )
teff_quant_w3 <- t(apply( X = teff_w3, MARGIN = 2, 
                     FUN = quantile, probs = c(0.1, 0.9) ))

teff_w4 <- tibble( school1 = post$`b_tXw4` + post$`r_school[1,tXw41]`,
                   school2 = post$`b_tXw4` + post$`r_school[2,tXw41]`,
                   school3 = post$`b_tXw4` + post$`r_school[3,tXw41]`,
                   school4 = post$`b_tXw4` + post$`r_school[4,tXw41]`,
                   school5 = post$`b_tXw4` + post$`r_school[5,tXw41]`,
                   school6 = post$`b_tXw4` + post$`r_school[6,tXw41]` )
teff_mean_w4 <- apply( X = teff_w4, MARGIN = 2, FUN = mean )
teff_quant_w4 <- t(apply( X = teff_w4, MARGIN = 2, 
                     FUN = quantile, probs = c(0.1, 0.9) ))

teff_w5 <- tibble( school1 = post$`b_tXw5` + post$`r_school[1,tXw51]`,
                   school2 = post$`b_tXw5` + post$`r_school[2,tXw51]`,
                   school3 = post$`b_tXw5` + post$`r_school[3,tXw51]`,
                   school4 = post$`b_tXw5` + post$`r_school[4,tXw51]`,
                   school5 = post$`b_tXw5` + post$`r_school[5,tXw51]`,
                   school6 = post$`b_tXw5` + post$`r_school[6,tXw51]` )
teff_mean_w5 <- apply( X = teff_w5, MARGIN = 2, FUN = mean )
teff_quant_w5 <- t(apply( X = teff_w5, MARGIN = 2, 
                     FUN = quantile, probs = c(0.1, 0.9) ))

su_teff <- tibble( school = rep(seq(1:6), 4), 
                   mean = c(teff_mean_w2, teff_mean_w3,
                            teff_mean_w4, teff_mean_w5), 
                   Q10 = c(teff_quant_w2[,1], teff_quant_w3[,1],
                           teff_quant_w4[,1], teff_quant_w5[,1]),
                   Q90 = c(teff_quant_w2[,2], teff_quant_w3[,2],
                           teff_quant_w4[,2], teff_quant_w5[,2]),
                   wave = rep(2:5, each=6) )
```
Graphical representation of results.

```{r}
gr_std_emostab_teff_across_schools <-
  ggplot(data = su_teff, aes(x = wave, y = mean)) +
  geom_pointrange(aes(ymin = Q10, 
                      ymax = Q90, group = school, color = school),
                  position=position_dodge(width=0.20)) +
  scale_x_discrete("Wave", limits=seq(2:6) )  +
  geom_hline(yintercept=0) + 
  theme(legend.position="bottom")
ggsave("../figures/std_emostab_teff_across_schools.pdf", gr_std_emostab_teff_across_schools, 
       width = 7, height = 4, units = "in")
gr_std_emostab_teff_across_schools
```




# Original computing environment

```{r, results='hide'}
devtools::session_info()
```

