---
title:  | 
        | Klasse im Puls
        | Model statement
author: 
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    fig_caption: yes
    theme: spacelab 
    highlight: pygments
    toc: TRUE
    toc_depth: 2
    number_sections: TRUE
    toc_float:
      smooth_scroll: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment=NA)
rm(list=ls()) # rm() needs to be specified; using list removes a bunch of objects; ls() lists all objects in the current workspace --> command removes all objects from R memory
library(tidyverse)
library(plm)       # Panel data analysis library
library(stargazer)
library(effects)
library(ggplot2)
library(brms)
set.seed(132435)
```


The purpose of this vignette is to develop and test a model framework for analyzing experiments in school classes.


# Data generating process 


Our aim is to estimate the effect of a classroom experiment. We consider one treatment group and one control group.


```{r}
class.size <- 30 # number of students in each class
student.id <- seq(1:class.size)
class.id <- seq(1:12)

wave <- seq(1:5) - 1

d <- expand.grid(wave = wave, student.id = student.id, class.id = class.id)

rm(wave, class.id, student.id)
```

First, we define the data structure for `r max(d$class.id)` groups (classrooms), `r class.size` students per class, and `r max(d$wave)+1` measurement occasions.

```{r}



d$treated <- ifelse(d$class.id %% 2 == 1, 1, 0)

  
  # ifelse(d$class.id==1 | d$class.id==3 | d$class.id==5 | d$class.id==7 | d$class.id==9 | d$class.id==11, 1, 0)


d$school.id <- ceiling(d$class.id / 2)


# d$school.id <- ifelse(d$class.id==1 | d$class.id==2, 1, ifelse(d$class.id==3 | d$class.id==4, 2, ifelse(d$class.id==5 | d$class.id==6, 3, ifelse(d$class.id==7 | d$class.id==8, 4, ifelse(d$class.id==9 | d$class.id==10, 5, ifelse(d$class.id==11 | d$class.id==12, 6,0))))))

df <- unique(d[c("student.id", "class.id")]) # creates new data sheet with those two variables from first data sheet
df <- df %>% mutate(ustudent.id = 1:nrow(df)) %>%
#mutate() adds new variables and preserves existing ones
    select(ustudent.id, everything()) #student ID für jeden einzelnen Schüler unabhängig von der Klasse
df$student.effect <- rnorm(nrow(df), mean=0, sd=1)
d <- left_join(d, df, by=c("class.id", "student.id"))
df <- unique(df <- unique(d[c("school.id")]))
df$school.effect <- rnorm(nrow(df), mean=0, sd=1.5)
d <- left_join(d, df, by="school.id")
tail(d)
```

The outcomes are determined by the following data generating process.

```{r}
teffect <- 1 # treatment effect
wave.effects <- -1 # wave effects
class.effect <- 2
cons <- 0
d$y <- cons + class.effect*d$treated + 
              wave.effects*d$wave + 
              teffect*(d$treated*d$wave) +
              d$student.effect +
              d$school.effect +
              rnorm(nrow(d), mean=0, sd=1)
```

# Difference in differences (DD)

```{r}
y00 <- mean( d$y[(d$treated==0 & d$wave==0)] )
y10 <- mean( d$y[(d$treated==1 & d$wave==0)] )
y01 <- mean( d$y[(d$treated==0 & d$wave==1)] )
y11 <- mean( d$y[(d$treated==1 & d$wave==1)] )
y02 <- mean( d$y[(d$treated==0 & d$wave==2)] )
y12 <- mean( d$y[(d$treated==1 & d$wave==2)] )
y03 <- mean( d$y[(d$treated==0 & d$wave==3)] )
y13 <- mean( d$y[(d$treated==1 & d$wave==3)] )
y04 <- mean( d$y[(d$treated==0 & d$wave==4)] )
y14 <- mean( d$y[(d$treated==1 & d$wave==4)] )

(y11 - y10) - (y01- y00) # (Veränderung in der Musikklasse) - (Veränderung in der Nicht-Musikklasse)
(y12 - y10) - (y02- y00)
(y13 - y10) - (y03- y00)
(y14 - y10) - (y04- y00)
```

# DD regression 

```{r}
d.p <- pdata.frame(d, index = c("ustudent.id", "wave"))
```

## OLS

```{r}
# olsall <- lm(y ~ treated + wave + treated*wave, data = d)
# efols <- effect("treated:wave", olsall)
# summary(efols)
# x <- as.data.frame(efols)
# ggplot(x, aes(treated, olsall, color=wave)) +
#   geom_point() + geom_errorbar(aes(ymin=olsall-se,
                                   # ymax=olsall+se), width=0.4) + theme_bw(base_size=12)

ols <- sapply(X=1:6, FUN=function(X){lm( y ~ treated + wave + treated*wave, data = d, subset = school.id==X)}, simplify = FALSE)

summary(ols[[1]])
summary(ols[[2]])
summary(ols[[3]])
summary(ols[[4]])
summary(ols[[5]])
summary(ols[[6]])
```

## Student fixed effects DD regression (within transformation)

```{r}
feWall <- plm(y ~ wave + treated*wave, data = d.p, model = "within")
summary(feWall)
```


```{r}
feW <- sapply(X=1:6, FUN=function(X){plm( y ~  wave + treated + treated*wave, data = d.p, model = "within", subset = school.id==X)}, simplify = FALSE)

summary(feW[[1]])
summary(feW[[2]])
summary(feW[[3]])
summary(feW[[4]])
summary(feW[[5]])
summary(feW[[6]])
```

## Student fixed effects DD regression (between transformation)

```{r}
feBall <- plm(y ~ wave + treated*wave, data = d.p)

feB <- sapply(X=1:6, FUN=function(X){plm( y ~  wave + treated + treated*wave, data = d.p, model = "between", subset = school.id==X)}, simplify = FALSE)

summary(feB[[1]])
summary(feB[[2]])
summary(feB[[3]])
summary(feB[[4]])
summary(feB[[5]])
summary(feB[[6]])
```

## Random effects DD regression

```{r}
re <- sapply(X=1:6, FUN=function(X){plm( y ~  treated + wave + treated*wave, data = d.p, model = "random", subset = school.id==X)}, simplify = FALSE)

summary(re[[1]])
summary(re[[2]])
summary(re[[3]])
summary(re[[4]])
summary(re[[5]])
summary(re[[6]])
```



New Analysis

```{r}
library(tidyverse)
library(Hmisc)
library(mosaic)
library(foreign)
library(lme4)
library(rstanarm)
```


# Model for six schools (treatments), two waves

## Model statment
We simultaneously investigate the treatments in six different schools. We restrict the analyses to wave 1 (pre treatment) and wave 2(post treatment).

```{r}
d$tXw1 <- as.factor( ifelse( (d$treated==1 & d$wave==1), 1, 0) )
d$tXw2 <- as.factor( ifelse( (d$treated==1 & d$wave==2), 1, 0) )
d$tXw3 <- as.factor( ifelse( (d$treated==1 & d$wave==3), 1, 0) )
d$tXw4 <- as.factor( ifelse( (d$treated==1 & d$wave==4), 1, 0) )
d2 <- d %>% filter( (wave==0 | wave==1)) %>% 
  select(school.id, ustudent.id, wave, treated, y, tXw1,)
table(d2$school.id, d2$treated)
```

## Fit with `lmer()`

```{r}
fit2 <- lmer( y ~ treated + wave + tXw1 + (1 + tXw1 | school.id) + (1 | ustudent.id), data = d2 )
summary(fit2)
```

## Fit with `rstanarm`

```{r}
bayes2 <- stan_lmer( y ~ treated + wave + tXw1 + (1 + treated + tXw1 | school.id) + (1 | ustudent.id), 
                data=d2, chains = 4, cores = 4, seed=123 )
summary(bayes2, 
        pars = c("(Intercept)", "treated1", "wave1", "tXw11",
                 "b[tXw11 school.id:1]", "b[tXw11 school.id:2]", "b[tXw11 school.id:3]",
                 "b[tXw11 school.id:4]", "b[tXw11 school.id:5]", "b[tXw11 school.id:6]"),
        probs = c(0.1, 0.5, 0.9),
        digits = 4)
```

## Posterior means

We extract the posterior draws to summarize the effect of the treatment.

```{r}
teff2 <- as.matrix(bayes2, 
                   pars = c("tXw11",
                            "b[tXw11 school.id:1]",
                            "b[tXw11 school.id:2]",
                            "b[tXw11 school.id:3]",
                            "b[tXw11 school.id:4]",
                            "b[tXw11 school.id:5]",
                            "b[tXw11 school.id:6]")) 
school.id1 <- teff2[,1] + teff2[,2]
school.id2 <- teff2[,1] + teff2[,3]
school.id3 <- teff2[,1] + teff2[,4] 
school.id4 <- teff2[,1] + teff2[,5] 
school.id5 <- teff2[,1] + teff2[,6] 
school.id6 <- teff2[,1] + teff2[,7] 
teff2 <- data.frame( school.id1 = school.id1,
                     school.id2 = school.id2,
                     school.id3 = school.id3,
                     school.id4 = school.id4,
                     school.id5 = school.id5,
                     school.id6 = school.id6)
head(teff2)
```

```{r}
r2_mean <- apply( X = teff2, MARGIN = 2, FUN = mean )
r2_quant <- apply( X = teff2, MARGIN = 2, 
                   FUN = quantile, probs = c(0.1, 0.9))
r2_quant <- data.frame(t(r2_quant))
names(r2_quant) <- c("Q10", "Q90")
r2 <- data.frame( school.id = seq(1:6), mean = r2_mean, r2_quant, wave = 1 )
r2 
```

Graphical representation of results.

```{r}
ggplot(data = r2, 
       aes(x = wave, 
           y = mean)) +
  geom_pointrange(aes(ymin = Q10, 
                      ymax = Q90, group = school.id, color = school.id),
                  position=position_dodge(width=0.20)) +
  scale_x_discrete("Wave", limits = c(2)) +
  geom_hline(yintercept=0)
```

# Full model for six schools and five waves

## Model statement

We simultaneously investigate the treatments in six different schools. We restrict the analysis to wave 0 (pre treatment) and wave 1 (post treatment).

```{r}
d3 <- d %>% select(school.id, ustudent.id, wave, treated, y, tXw1, tXw2, tXw3, tXw4)
table(d3$school.id, d3$treated)
```

## Fit with `lmer()`

```{r}
fit3 <- lmer( y ~ treated + wave +  tXw1 + tXw2 + tXw3 + tXw4 + 
                (1 + treated + tXw1 + tXw2 + tXw3 + tXw4 | school.id) +
                       (1 | ustudent.id), data=d3)
coef(fit3)$school.id
```

## Fit with `rstanarm`

```{r, message=FALSE, cache=TRUE}
bayes3 <- stan_lmer( y ~ treated + wave + tXw1 +tXw2 + tXw3 + tXw4 +
                       (1 + treated + tXw1 + tXw2 + tXw3 + tXw4 | school.id) +
                       (1 | ustudent.id), 
                data=d3, chains = 4, cores = 4, seed=123)
summary(bayes3, 
        pars = c("(Intercept)", "treated1", "welle1", "welle2", "welle3", "welle4",
                 "tXw51", "tXw21", "tXw31", "tXw41"),
        probs = c(0.1, 0.5, 0.9), digits = 4)
```

## Posterior means

We extract the posterior draws to summarize the effect of the treatment.

```{r}
summary(bayes3, 
        pars = c("tXw31", 
                 "b[tXw31 school:1]", "b[tXw31 school:2]", "b[tXw31 school:3]", 
                 "b[tXw31 school:4]", "b[tXw31 school:5]", "b[tXw31 school:6]"),
        probs = c(0.1, 0.5, 0.9), digits = 4)
```

```{r}
teff5 <- as.matrix(bayes3, 
                   pars = c("tXw11",
                            "b[tXw11 school.id:1]",
                            "b[tXw11 school.id:2]",
                            "b[tXw11 school.id:3]",
                            "b[tXw11 school.id:4]",
                            "b[tXw11 school.id:5]",
                            "b[tXw11 school.id:6]",
                            "tXw21",
                            "b[tXw21 school.id:1]",
                            "b[tXw21 school.id:2]",
                            "b[tXw21 school.id:3]",
                            "b[tXw21 school.id:4]",
                            "b[tXw21 school.id:5]",
                            "b[tXw21 school.id:6]",
                            "tXw31",
                            "b[tXw31 school.id:1]",
                            "b[tXw31 school.id:2]",
                            "b[tXw31 school.id:3]",
                            "b[tXw31 school.id:4]",
                            "b[tXw31 school.id:5]",
                            "b[tXw31 school.id:6]",
                            "tXw41",
                            "b[tXw41 school.id:1]",
                            "b[tXw41 school.id:2]",
                            "b[tXw41 school.id:3]",
                            "b[tXw41 school.id:4]",
                            "b[tXw41 school.id:5]",
                            "b[tXw41 school.id:6]")) 

wave1.school.id1 <- teff5[,"tXw11"] + teff5[,"b[tXw11 school.id:1]"]
wave1.school.id2 <- teff5[,"tXw11"] + teff5[,"b[tXw11 school.id:2]"]
wave1.school.id3 <- teff5[,"tXw11"] + teff5[,"b[tXw11 school.id:3]"]
wave1.school.id4 <- teff5[,"tXw11"] + teff5[,"b[tXw11 school.id:4]"]
wave1.school.id5 <- teff5[,"tXw11"] + teff5[,"b[tXw11 school.id:5]"]
wave1.school.id6 <- teff5[,"tXw11"] + teff5[,"b[tXw11 school.id:6]"]
wave2.school.id1 <- teff5[,"tXw21"] + teff5[,"b[tXw21 school.id:1]"]
wave2.school.id2 <- teff5[,"tXw21"] + teff5[,"b[tXw21 school.id:2]"]
wave2.school.id3 <- teff5[,"tXw21"] + teff5[,"b[tXw21 school.id:3]"]
wave2.school.id4 <- teff5[,"tXw21"] + teff5[,"b[tXw21 school.id:4]"]
wave2.school.id5 <- teff5[,"tXw21"] + teff5[,"b[tXw21 school.id:5]"]
wave2.school.id6 <- teff5[,"tXw21"] + teff5[,"b[tXw21 school.id:6]"]
wave3.school.id1 <- teff5[,"tXw31"] + teff5[,"b[tXw31 school.id:1]"]
wave3.school.id2 <- teff5[,"tXw31"] + teff5[,"b[tXw31 school.id:2]"]
wave3.school.id3 <- teff5[,"tXw31"] + teff5[,"b[tXw31 school.id:3]"]
wave3.school.id4 <- teff5[,"tXw31"] + teff5[,"b[tXw31 school.id:4]"]
wave3.school.id5 <- teff5[,"tXw31"] + teff5[,"b[tXw31 school.id:5]"]
wave3.school.id6 <- teff5[,"tXw31"] + teff5[,"b[tXw31 school.id:6]"]
wave4.school.id1 <- teff5[,"tXw41"] + teff5[,"b[tXw41 school.id:1]"]
wave4.school.id2 <- teff5[,"tXw41"] + teff5[,"b[tXw41 school.id:2]"]
wave4.school.id3 <- teff5[,"tXw41"] + teff5[,"b[tXw41 school.id:3]"]
wave4.school.id4 <- teff5[,"tXw41"] + teff5[,"b[tXw41 school.id:4]"]
wave4.school.id5 <- teff5[,"tXw41"] + teff5[,"b[tXw41 school.id:5]"]
wave4.school.id6 <- teff5[,"tXw41"] + teff5[,"b[tXw41 school.id:6]"]

wave1 <- data.frame( school.id1 = wave1.school.id1,
                     school.id2 = wave1.school.id2,
                     school.id3 = wave1.school.id3,
                     school.id4 = wave1.school.id4,
                     school.id5 = wave1.school.id5,
                     school.id6 = wave1.school.id6)
wave2 <- data.frame( school.id1 = wave2.school.id1,
                     school.id2 = wave2.school.id2,
                     school.id3 = wave2.school.id3,
                     school.id4 = wave2.school.id4,
                     school.id5 = wave2.school.id5,
                     school.id6 = wave2.school.id6)
wave3 <- data.frame( school.id1 = wave3.school.id1,
                     school.id2 = wave3.school.id2,
                     school.id3 = wave3.school.id3,
                     school.id4 = wave3.school.id4,
                     school.id5 = wave3.school.id5,
                     school.id6 = wave3.school.id6)
wave4 <- data.frame( school.id1 = wave4.school.id1,
                     school.id2 = wave4.school.id2,
                     school.id3 = wave4.school.id3,
                     school.id4 = wave4.school.id4,
                     school.id5 = wave4.school.id5,
                     school.id6 = wave4.school.id6)

```

```{r}
wave1_mean <- apply( X = wave1, MARGIN = 2, FUN = mean )
wave2_mean <- apply( X = wave2, MARGIN = 2, FUN = mean )
wave3_mean <- apply( X = wave3, MARGIN = 2, FUN = mean )
wave4_mean <- apply( X = wave4, MARGIN = 2, FUN = mean )

wave1_quant <- apply( X = wave1, MARGIN = 2, 
                   FUN = quantile, probs = c(0.25, 0.75))
wave2_quant <- apply( X = wave2, MARGIN = 2, 
                   FUN = quantile, probs = c(0.25, 0.75))
wave3_quant <- apply( X = wave3, MARGIN = 2, 
                   FUN = quantile, probs = c(0.25, 0.75))
wave4_quant <- apply( X = wave4, MARGIN = 2, 
                   FUN = quantile, probs = c(0.25, 0.75))

wave1_quant <- data.frame(t(wave1_quant))
wave2_quant <- data.frame(t(wave2_quant))
wave3_quant <- data.frame(t(wave3_quant))
wave4_quant <- data.frame(t(wave4_quant))

names(wave1_quant) <- c("Q10", "Q90")
names(wave2_quant) <- c("Q10", "Q90")
names(wave3_quant) <- c("Q10", "Q90")
names(wave4_quant) <- c("Q10", "Q90")

r.wave1 <- data.frame( school.id = seq(1:6), mean = wave1_mean, wave1_quant, wave = 5 )
r.wave2 <- data.frame( school.id = seq(1:6), mean = wave2_mean, wave2_quant, wave = 2 )
r.wave3 <- data.frame( school.id = seq(1:6), mean = wave3_mean, wave3_quant, wave = 3 )
r.wave4 <- data.frame( school.id = seq(1:6), mean = wave4_mean, wave4_quant, wave = 4 )
r.full <- rbind.data.frame(r.wave1, r.wave2, r.wave3, r.wave4)
```

Graphical representation of results.

```{r}
ggplot(data = r.full, 
       aes(x = wave, 
           y = mean)) +
  geom_pointrange(aes(ymin = Q10, 
                      ymax = Q90, group = school.id, color = school.id),
                  position=position_dodge(width=0.50)) +
  scale_x_discrete("Wave", limits = c(2,3,4,5) ) +
  geom_hline(yintercept=0)
```

# Original computing environment

```{r}
devtools::session_info()
```

